Demystifying the Computer Science PhD Admission in the US\\{\large A Guide for International Students

### 2023-11-19 demystify.tex
updates

\usepackage[T1]{fontenc}
\usepackage[margin=1.5in]{geometry}
\usepackage{soul}
%\usepackage[small,compact]{titlesec} %very powerful
\usepackage[most]{tcolorbox}
\usepackage{enumitem}
\usepackage{epigraph}
\usepackage{cite}
\usepackage{caption}
\captionsetup{font=small}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{wrapfig}
\setlength\intextsep{0pt} % remove extra space above and below in-line float
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=blue,
    urlcolor=blue,
}
\usepackage{booktabs}

@@

\newenvironment{commentbox}[1][]{
\small
    \begin{mybox}
    {\small \textbf{#1}}
 }{
   \end{mybox}
}
\renewcommand{\figurename}{Fig.}
\renewcommand{\tablename}{Tab.}
\def\Section{\S}
\newcommand{\mycomment}[3][\color{blue}]{{#1{{#2}: {#3}}}}
\newcommand{\tvn}[1]{\mycomment{TVN}{#1}}{}
\newcommand{\didi}[1]{\mycomment{Didier}{#1}}{}
\newcommand{\tl}[1]{\mycomment{ThanhLe}{#1}}{}
\newcommand{\red}[1]{{\color{red}{#1}}}
\newcommand{\xz}[1]{\mycomment{Xiaokuan}{[#1]}}{}

@@

@@

\begin{center}
\includegraphics[width=0.6\linewidth]{c6.png}
\end{center}

@@

    \begin{center}
      \includegraphics[scale=0.5]{c2.png}
    \end{center}

@@

@@

\subsection{Selecting and Ranking Schools for CS}
   \begin{center}
    \includegraphics[scale=0.5]{c1.png}
   \end{center}

@@

Here are some general advice to find professor.  First, search for profs. that share similar research interests. You can learn about faculty and their level of research activity through the faculty's website or even from CSRankings.  For example, in CSRankings, if you want to work with PL, you can search for those publish in PL conferences.  If you want to work with SE \emph{and} AI, you can search for faculty who work in both SE and AI.   You can also go to their website, look at their papers and projects.  In any case, you should contact the professor and talk to them. For example, you can (cold) email them and ask about their research and if they are taking new students and schedule appointments to chat with them (office hrs are often the best time).  See \S\ref{sec:contact} for more details on contacting professors.

Another very effective way is taking graduate level courses in the topics you are interested in (remember: you have those 2 years to explore).  Professors teaching these \emph{special topics courses} and \emph{research seminars} typically are recruiting students---giving you even a higher chance. Do well in the class, answer questions, talk to the profs after classes, etc---being stand out.  Many professors, including myself, prefer taking in new students this way.  It gives both the professor and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds a bit like a marriage!).  
You can also ask if you can do an independent study or research with a prof. This can be informal (no credit) and takes place during the summer or winter break.  For example, I work with multiple students, many of whom are undergrads, this way. Many will drop because they find they don't like my research, but some will stay and become my PhD students.

@@

      \begin{center}
        \includegraphics[scale=0.4]{c8.png}
      \end{center}

@@

Faculty received many "cold" e-mails from international students seeking for admission, TA, and RA. Most of the time, we ignore these emails, but in some rare occasions we do answer them. So how to write an email that can get our attention?

@@

\begin{center}
    \includegraphics[scale=0.2]{alpinist-climbing-peak-mountain-comic-hand-drawn-vector-illustration.jpg}
  \end{center}

@@

\begin{center}
\includegraphics[scale=0.3]{c4a.png}
\end{center}

@@

\begin{center}
\includegraphics[scale=0.6]{c5.png}
\end{center}

@@

      \begin{center}
        \includegraphics[scale=0.3]{c6.png}
      \end{center}

@@

>>>>

\usepackage[T1]{fontenc}
%\usepackage[margin=1.5in]{geometry}
\usepackage{soul}
%\usepackage[small,compact]{titlesec} %very powerful
\usepackage[most]{tcolorbox}
\usepackage{enumitem}
\usepackage{epigraph}
\usepackage{cite}
\usepackage{caption}
\captionsetup{font=small}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{wrapfig}
\setlength\intextsep{0pt} % remove extra space above and below in-line float
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=blue,
    urlcolor=blue,
}
\usepackage{booktabs}

@@

\newenvironment{commentbox}[1][]{
\small
    \begin{mybox}
    {\small \textbf{#1}}
 }{
   \end{mybox}
}

\newtcolorbox{mydomesticbox}{
enhanced,
boxrule=0pt,frame hidden,
borderline west={2pt}{0pt}{red!75!black},
colback=blue!10!white,
sharp corners
}

\newenvironment{domesticbox}[1][]{
\small
    \begin{mydomesticbox}
    {\small \textbf{#1}}
 }{
   \end{mydomesticbox}
}

\renewcommand{\figurename}{Fig.}
\renewcommand{\tablename}{Tab.}
\def\Section{\S}
\newcommand{\mycomment}[3][\color{blue}]{{#1{{#2}: {#3}}}}
\newcommand{\tvn}[1]{\mycomment{TVN}{#1}}{}
\newcommand{\didi}[1]{\mycomment{Didier}{#1}}{}
\newcommand{\tl}[1]{\mycomment{ThanhLe}{#1}}{}
\newcommand{\red}[1]{{\color{red}{#1}}}
\newcommand{\xz}[1]{\mycomment{Xiaokuan}{[#1]}}{}

@@

\begin{domesticbox}[Domestic students:] If you're a domestic student, you actually have several advantages in your application.  First, we already know about your school (see \S\ref{sec:your-school}), and you are also more familiar with the US education system and culture.  You also have more opportunities for funding, e.g., through government scholarships for US citizens and residents. Finally, compared to international students, PhD applications in CS from domestic students are also few and that can help your case.
\end{domesticbox}

@@

\begin{center}
\includegraphics[width=0.6\linewidth]{files/c6.png}
\end{center}

@@

    \begin{center}
      \includegraphics[scale=0.5]{files/c2.png}
    \end{center}

@@

\paragraph{If you do not get admitted} If you do not get admitted to any schools (or don't want to go to the ones that you got it), try again next time.  Graduate admission can involve randomness and noise.  Don't bother asking for feedback, you will not likely get any.  In the meantime, you can work on improving your profile, e.g., publish more papers, get more research experiences, etc.

You can also consider applying to MS programs, which are typically easier to get in (but you need to pay).  If you get in an MS at a school of your choice, you can contact professors to work with them. If you do well, you can ask the professor to support you to convert to PhD.  This is a common path for many students.

@@

\subsection{Selecting and Ranking Schools for CS}
   \begin{center}
    \includegraphics[scale=0.5]{files/c1.png}
   \end{center}

@@

Here are some general advice to find professor.  First, search for profs. that share similar research interests. You can learn about faculty and their level of research activity through the faculty's website or even from CSRankings.  For example, in CSRankings, if you want to work with PL, you can search for those publish in PL conferences.  If you want to work with SE \emph{and} AI, you can search for faculty who work in both SE and AI.   You can also go to their website, look at their papers and projects.  In any case, you should contact the professor and talk to them. For example, you can (cold) email them and ask about their research and if they are taking new students (\S\ref{sec:contact}) and schedule appointments to chat with them (office hrs are often the best time).

Another effective way is taking graduate level courses in the topics you are interested in (remember: you have those 2 years to explore).  Professors teaching these \emph{special topics courses} and \emph{research seminars} typically are recruiting students---giving you even a higher chance. Do well in the class, answer questions, talk to the profs after classes, etc---being stand out.  Many professors, including myself, prefer taking in new students this way.  It gives both the professor and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds a bit like a marriage!).
You can also ask if you can do an independent study or research with a prof. This can be informal (no credit) and takes place during the summer or winter break.  For example, I work with multiple students, many of whom are undergrads, this way. Many will drop because they find they don't like my research, but some will stay and become my PhD students.

@@

      \begin{center}
        \includegraphics[scale=0.4]{files/c8.png}
      \end{center}

@@

\begin{center}
\includegraphics[scale=0.5]{files/emails.png}
\end{center}

Faculty received many "cold" e-mails from international students seeking for admission, TA, and RA. Most of the time, we ignore these emails, but in some rare occasions we do answer them. So why do we ignore some but reply to others and how to write an email that get our attention?

@@

\begin{center}
    \includegraphics[scale=0.2]{files/alpinist-climbing-peak-mountain-comic-hand-drawn-vector-illustration.jpg}
  \end{center}

@@

\begin{center}
\includegraphics[scale=0.3]{files/c4a.png}
\end{center}

@@

\begin{center}
\includegraphics[scale=0.6]{files/c5.png}
\end{center}

@@

      \begin{center}
        \includegraphics[scale=0.3]{files/c6.png}
      \end{center}

@@

\chapter{References}
